---
title: "Lab 10"
author: "Corinna Hong"
date: "March 12, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}

library(tidyverse)
library(here)
library(DBI)
library(RSQLite)

```


### Connect to database and tell R we are using SQLite

```{r}

lter_db <- DBI::dbConnect(drv = RSQLite::SQLite(),
                          dbname = here("data","marine.sqlite")) # Use SQLite engine as our driver and connect to marine.sqlite database

```


### What tables existin the database?
To check what tables exist by the names, use `DBI::dbListTables()`

```{r}

DBI::dbListTables(lter_db)

```


### Which varables are in a database table?
Use `DBI::dbListFields()`

```{r}

DBI::dbListFields(lter_db, "fish_data") # What are cols in fish data?
DBI::dbListFields(lter_db, "inverts_data")
DBI::dbListFields(lter_db, "lobsters_data")

```


### Read in database table as an R data frame
Ues `DBI::dbReadTable()`

```{r}

fish_df <- DBI::dbReadTable(lter_db, "fish_data") # It is not an R data frame!

```


### Query a database using SQL
- SELECT: gets columns
- FROM: what table in a database to get something from
- Use * to indicate "get all"

```{sql connection=lter_db}

SELECT * FROM fish_data

```

To get a subset of columns, we can just list them after SELECT, separated by a comma

```{sql connection=lter_db}

SELECT year, site FROM fish_data

```

To get ROWS (observations) that match a given condition we set, use WHERE

```{sql connection=lter_db}

SELECT * FROM fish_data WHERE year == 2016

```

Query to keep columns, but only if the common_name variable matches "garibaldi"

```{sql connection=lter_db}

SELECT common_name, site FROM fish_data WHERE common_name == "garibaldi"

```

Query to keep observations where common_name == "garibaldi" AND site == "mohk"

```{sql connection=lter_db}

SELECT * FROM fish_data WHERE common_name == "garibaldi" AND site == "mohk"

```

Query to keep obervations where year is 2017 OR site is "abur"

```{sql connection=lter_db}

SELECT * FROM fish_data WHERE year == "2017" OR site == "abur"

```


### Store an output for SQL Querying
Add in code chunk header, `output.var = "name_object"`

```{sql connection=lter_db, output.var = "fish_2017_df"}

SELECT * FROM fish_data WHERE year == 2017

```


### Query using SQL from an R code chunk
`DBI::dbSendQuery()`

```{r}

garibaldi <- dbSendQuery(lter_db, "SELECT * FROM fish_data WHERE common_name == 'garibaldi'") # put SQL code in quotations, use single quotations so R doesn't get mad with nested double quotations

gar_df <- dbFetch(garibaldi) # Turn it into a df!

```


### Querying using dplyr-ish code
`dbplyr` don;t need to tall it, R knows to use it

```{r}

fish_tbl <- tbl(lter_db, "fish_data")


# Wrangle the table
yearly_fish <- fish_tbl %>% 
  group_by(year) %>% 
  summarize(
    total_fish = sum(total_count)
  )

yearly_fish %>% show_query() # Tells you what SQL code you need to input into an SQL chunk to get it to do the same thing as the above dbplyr code

yearly_fish %>% collect() # Gives you outcome of SQL code

```


### Update a SQL database
Ues `DBI::dbWriteTable` and `DBI::dbRemoveTable` functions

```{r}

DBI::dbWriteTable(lter_db, "mtcars_df", mtcars) # tell it which db to add it to, what to name it, and what R calls it

# Did it work?
DBI::dbListTables(lter_db)

# Remove it
DBI::dbRemoveTable(lter_db, "mtcars_df")

```







